Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-19 ind_files/01_Introduction.tex
Add some references

The most general mathematical framework we have for evolutionary population biology is the Price equation. Indeed, much like statistical mechanics in physics, the Price equation is derived from a small number of very general first principles and is able to recover several standard equations as special cases. The Price equation partitions changes in population composition into multiple terms, each of which lends itself to a straightforward interpretation in terms of `high-level' evolutionary forces such as selection and mutation, thus providing a useful conceptual framework for thinking about how populations change over time~\citep{frank_natural_2012}. However, the greater complexity of ecology and evolution relative to physics has meant that the generality of the Price equation comes with a much bigger cost in predictive power - the Price equation in its most general formulation is dynamically insufficient\footnote{Since this is not very standard nomenclature outside theoretical biology and related fields: `Dynamically insufficient' means that the equation cannot be iterated, or presented in the form of a difference equation $x_{t+1} = F(x_t)$ or a differential equation $\dot{x} = F(x)$. Thus, we cannot predict a dynamic `trajectory' given an initial condition $x_0$. The term `insufficient' is to indicate that the equation requires `complete' information to be true and only relates two quantities `in retrospect'. It is thus `insufficient' for prediction. In our case, in its most general setting, the Price equation is usually formulated in a way that partitions a \emph{given} amount of phenotypic change between two populations (usually but not necessarily the same population at two different times) into change due to selection, transmission bias, etc., rather than \emph{predicting} a trajectory for how much phenotypic change will occur at various future times based on the \emph{current} phenotypic distribution.}~\citep{van_veelen_use_2005,frank_natural_2012,luque_one_2017}. However, this need not be the case, and indeed, several authors have put forth more predictive versions of the Price equation by moving to a continuous time differential equation framework in which the Price equation is dynamically sufficient but manifests in a slightly less general form~\citep{page_unifying_2002,lion_theoretical_2018}.

@@

One of the general guiding principles of much of this mathematization has been the assumption that incorporating the reality of finite population sizes into models leads to no major qualitative differences in behavior, only `adding noise' or `blurring out' the predictions of simpler infinite population models~\citep{page_unifying_2002}. Consequently, several major theoretical frameworks in the field, such as adaptive dynamics, are explicitly formulated in deterministic terms at the infinite population size limit. However, this assumption is largely unjustified, and since populations in the real world are finite and stochastic, checking whether stochastic models differ from their deterministic analogs is vital to furthering our understanding of the fundamentals of population biology~\citep{hastings_transients_2004, coulson_skeletons_2004, shoemaker_integrating_2020}. Today, we increasingly recognize that incorporating the finite and stochastic nature of the real world routinely has much stronger consequences than simply `adding noise' to deterministic expectations~\citep{boettiger_noise_2018}, with important consequences for both ecological~\citep{schreiber_does_2022} and evolutionary~\citep{delong_stochasticity_2023} theory. In ecology and evolution, stochastic models need not exhibit phenomena predicted by their deterministic analogues~\citep{proulx_what_2005, johansson_will_2006, claessen_delayed_2007,  wakano_evolutionary_2013, debarre_evolutionary_2016, johnson_two-dimensional_2021}. In addition, they exhibit novel phenomena not predicted by the deterministic approximations~\citep{rogers_demographic_2012, rogers_spontaneous_2012, rogers_modes_2015, veller_drift-induced_2017, delong_stochasticity_2023}, for example, even completely `reversing' the predictions of deterministic models~\citep{houchmandzadeh_selection_2012,houchmandzadeh_fluctuation_2015,constable_demographic_2016,mcleod_social_2019}..

Studies of neutral or near-neutral dynamics in population and quantitative genetics usually do take stochasticity seriously, explicitly modeling finite populations that follow stochastic dynamics. Unfortunately, the classic or standard stochastic models in both population genetics~\citep{fisher_genetical_1930,wright_evolution_1931, moran_random_1958, kimura_diffusion_1964} and quantitative genetics~\citep{crow_introduction_1970, lande_natural_1976} typically assume a fixed total population size, thus restricting their validity in a world where population sizes routinely fluctuate. Those models which study evolution in populations with non-constant size usually impose deterministic and typically phenomenological rules for how the total population size must vary~\citep{kimura_probability_1974,ewens_probability_1967,otto_probability_1997, engen_fixation_2009, waxman_unified_2011}. These rules further usually do not depend on population composition. Such models are thus somewhat artificial since demography and population size are forced to be independent quantities even though this is obviously not the case in natural populations, where population size is a `bulk' property whose value emerges from an intricate interplay of the individual-level demographic processes of birth and death~\citep{metcalf_why_2007,doebeli_towards_2017}. Notably, as mentioned above, the most general framework we have, the Price equation, is usually formulated as a deterministic, dynamically insufficient equation (but see~\cite{page_unifying_2002} and~\cite{lion_theoretical_2018} for a dynamically sufficient formulation that is deterministic and~\cite{rice_stochastic_2008} for a stochastic formulation that is dynamically insufficient). Since real-life populations are stochastic, finite, and of non-constant population size, this is somewhat of a problem, since we know that such deterministic approximations that may not capture important dynamics of the real systems of interest.

>>>>

The most general mathematical framework we have for evolutionary population biology is the Price equation. Indeed, much like statistical mechanics in physics, the Price equation is derived from a small number of very general first principles and is able to recover several standard equations as special cases. The Price equation partitions changes in population composition into multiple terms, each of which lends itself to a straightforward interpretation in terms of `high-level' evolutionary forces such as selection and mutation, thus providing a useful conceptual framework for thinking about how populations change over time~\citep{frank_natural_2012}. However, the greater complexity of ecology and evolution relative to physics has meant that the generality of the Price equation comes with a much bigger cost in predictive power - the Price equation in its most general formulation is dynamically insufficient\footnote{Since this is not very standard nomenclature outside theoretical biology and related fields: `Dynamically insufficient' means that the equation cannot be iterated, or presented in the form of a difference equation $x_{t+1} = F(x_t)$ or a differential equation $\dot{x} = F(x)$. Thus, we cannot predict a dynamic `trajectory' given an initial condition $x_0$. The term `insufficient' is to indicate that the equation requires `complete' information to be true and only relates two quantities `in retrospect'. It is thus `insufficient' for prediction. In our case, in its most general setting, the Price equation is usually formulated in a way that partitions a \emph{given} amount of phenotypic change between two populations (usually but not necessarily the same population at two different times) into change due to selection, transmission bias, etc., rather than \emph{predicting} a trajectory for how much phenotypic change will occur at various future times based on the \emph{current} phenotypic distribution.}~\citep{van_veelen_use_2005,frank_natural_2012,luque_one_2017}. However, this need not be the case, and indeed, several authors have put forth more predictive versions of the Price equation by moving to a continuous time differential equation framework in which the Price equation is dynamically sufficient but manifests in a slightly less general form~\citep{page_unifying_2002,lion_theoretical_2018,day_price_2020}.

@@

One of the general guiding principles of much of this mathematization has been the assumption that incorporating the reality of finite population sizes into models leads to no major qualitative differences in behavior, only `adding noise' or `blurring out' the predictions of simpler infinite population models~\citep{page_unifying_2002}. Consequently, several major theoretical frameworks in the field, such as adaptive dynamics, are explicitly formulated in deterministic terms at the infinite population size limit. However, this assumption is largely unjustified, and since populations in the real world are finite and stochastic, checking whether stochastic models differ from their deterministic analogs is vital to furthering our understanding of the fundamentals of population biology~\citep{hastings_transients_2004, coulson_skeletons_2004, shoemaker_integrating_2020}. Today, we increasingly recognize that incorporating the finite and stochastic nature of the real world routinely has much stronger consequences than simply `adding noise' to deterministic expectations~\citep{boettiger_noise_2018}, with important consequences for both ecological~\citep{schreiber_does_2022} and evolutionary~\citep{delong_stochasticity_2023} theory. In ecology and evolution, stochastic models need not exhibit phenomena predicted by their deterministic analogues~\citep{proulx_what_2005, johansson_will_2006, claessen_delayed_2007,  wakano_evolutionary_2013, debarre_evolutionary_2016, johnson_two-dimensional_2021}. In addition, they exhibit novel phenomena not predicted by the deterministic approximations~\citep{rogers_demographic_2012, rogers_spontaneous_2012, rogers_modes_2015, veller_drift-induced_2017, delong_stochasticity_2023}, for example, even completely `reversing' the predictions of deterministic models~\citep{houchmandzadeh_selection_2012,houchmandzadeh_fluctuation_2015,constable_demographic_2016,mcleod_social_2019}.

Studies of neutral or near-neutral dynamics in population and quantitative genetics usually do take stochasticity seriously, explicitly modeling finite populations that follow stochastic dynamics. Unfortunately, the classic or standard stochastic models in both population genetics~\citep{fisher_genetical_1930,wright_evolution_1931, moran_random_1958, kimura_diffusion_1964} and quantitative genetics~\citep{crow_introduction_1970, lande_natural_1976} typically assume a fixed total population size, thus restricting their validity in a world where population sizes routinely fluctuate. Those models which study evolution in populations with non-constant size usually impose deterministic and typically phenomenological rules for how the total population size must vary~\citep{kimura_probability_1974,ewens_probability_1967,otto_probability_1997, engen_fixation_2009, waxman_unified_2011}. These rules further usually do not depend on population composition. Such models are thus somewhat artificial since demography and population size are forced to be independent quantities even though this is obviously not the case in natural populations, where population size is a `bulk' property whose value emerges from an intricate interplay of the individual-level demographic processes of birth and death~\citep{metcalf_why_2007,doebeli_towards_2017}. Notably, the most general framework we have, the Price equation, is typically formulated in a deterministic setting ~\citep{page_unifying_2002,frank_natural_2012,queller_fundamental_2017,lion_theoretical_2018,day_price_2020} that ignores stochasticity (but see~\cite{rice_stochastic_2008,rice_universal_2020} for a discrete time formulation that is stochastic but, as far as I can tell, is dynamically insufficient, just like the original formulation of the Price equation). Since real-life populations are stochastic, finite, and of non-constant population size, this is somewhat of a problem, since we know that such deterministic approximations that may not capture important dynamics of the real systems of interest.

### 2023-03-19 ind_files/06_nontechnical_discussion.tex
Add some references

The equations of section \ref{sec_fun_theorems} generically predict a directional evolutionary force acting on variation in per-capita turnover rate $\tau$ that I call `noise-induced selection' (following previous studies such as~\cite{constable_demographic_2016} and~\cite{week_white_2021}). Noise-induced selection is only seen in finite populations, is seen whenever there is differential turnover rate $\tau$ in the system, and arises due to different types of individuals in the population experiencing a different number of stochastic events (birth and death) in a given time interval. Several specific finite population models, especially in social evolution, illustrate that evolution in finite population can proceed in a direction different from that predicted by the infinite population limit~\citep{parsons_consequences_2010,melbinger_evolutionary_2010, houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,debarre_evolutionary_2016, behar_fluctuations-induced_2016, constable_demographic_2016,abu_awad_effects_2018,parsons_pathogen_2018,mcavoy_public_2018,mcleod_social_2019}. Indeed, in some stochastic models, the outcome of finite population models can be exactly opposite to that of the infinite population limit, a phenomenon sometimes referred to as `reversing the direction of deterministic selection'~\citep{constable_demographic_2016,mcleod_social_2019}. Chapter \ref{chap_unification} explains how the  results of all these studies can be explained using noise-induced selection. Most recently, a study showed that in a very broad class of competition models of population ecology, finite populations can directionally evolve in a manner opposite to the direction of evolution predicted in the infinite population limit, essentially due to the effects of noise-induced selection~\citep{mazzolini_universality_2022}. My thesis generalizes this particular result of~\cite{mazzolini_universality_2022} to models with arbitrary interaction types and presents the relevant equations in a form that highlights connections with standard results in evolutionary theory such as the Price equation. It is worth noting that the existence of noise-induced selection directly implies (from equation \eqref{nD_stochastic_RM} or \eqref{nD_stochastic_Price}) that evolution is not expected to maximize fitness in finite populations even if fitness is entirely frequency independent as long as there is some (heritable) variation in the turnover rates $\tau_i$, further underscoring the now well-appreciated point that the view of evolution as `climbing a hill' on a fitness landscape and thereby maximizing fitness is rather limited~\citep{grodwohl_theory_2017}.

The equations of section \ref{sec_fun_theorems} also imply that for the evolution of a trait to be \emph{truly} neutral in finite populations (in the sense of all $m$ types in a system having equal fixation/extinction probability if we start with an initial state in which every type has frequency $1/m$), it is not sufficient for the trait in question to be neutral with respect to fitness $w$. Instead, we also require the trait to be neutral with respect to turnover rate $\tau$.  Indeed, in ecological models, previous work has numerically shown that in finite, fluctuating populations, the equal growth rate of types is not sufficient to ensure equal fixation probabilities and that there is a slight biasing for types with lower turnover rates, sometimes interpreted as a selection `for longevity'~\citep{lin_features_2012, oliveira_advantage_2017,balasekaran_quasi-neutral_2022}. In models of evolutionary game theory in fluctuating finite populations, individuals with lower death rates have higher fixation probability even when growth rates are equalized~\citep{huang_stochastic_2015, czuppon_fixation_2018}. Similarly, models of cell cycle dynamics find that selection favors cell types that periodically arrest their cell cycle relative to non-arresting cells even when their growth rates are equal~\citep{wodarz_effect_2017}. In the language of the birth-death formalism I develop in this thesis, all of these studies equalize the growth rates $w$ of competing types but allow the turnover to vary (by reducing the birth rate through arresting the cell cycle, for example), thus allowing noise-induced selection for reduced turnover to operate in the system. My derivations show analytically that such deviations from neutrality in finite populations are a generic phenomenon explained by noise-induced selection and should be expected whenever there is variation in turnover rates. Selection on turnover rates also leads to insights on life-history evolution, and these insights have been extensively reported in a recent pre-print that independently arrives at my equations for type frequencies (equation \eqref{nD_stochastic_RM}) and the change of mean fitness and turnover in the population (equations \eqref{nD_stochastic_Fisher} and \eqref{nD_stochastic_Fisher_turnover} respectively) using certain discrete time stochastic processes and their approximation via techniques reminiscent of numerical stochastic integration~\citep{kuosmanen_turnover_2022}.

On the practical side, the existence of noise-induced selection implies that simulation studies working with evolutionary individual-based or agent-based models should be careful about whether interaction effects are incorporated into birth rates or death rates since this seemingly arbitrary choice can have unintended consequences due to noise-induced selection, thus potentially biasing results~\citep{mcleod_social_2019,kuosmanen_turnover_2022}. My results also indicate that measuring the growth rate of populations is not, in general, sufficient for accurate prediction/inference of future trajectories of the relative abundance of a species (or phenotype, allele, etc.) from empirical data even in completely controlled environments. The growth rate $w_i = b^{\textrm{(ind)}}_{i} - d^{\textrm{(ind)}}_{i}$ of a species $i$ only specifies the difference between its per-capita birth and death rates. In contrast, the complete stochastic dynamics also depend on the total turnover $\tau_i = b^{\textrm{(ind)}}_{i} + d^{\textrm{(ind)}}_{i}$ (\emph{i.e.} the sum of the per-capita birth and death rates).

@@

Lastly, noise-induced selection is particular to fluctuating populations and does not occur in models with fixed population sizes\footnote{If $\sum_j x_j$ is a constant, the map $x_i \to x_i/\sum_j x_j$ becomes a linear map and we no longer need It\^o's formula to move from densities to frequencies in the derivation I conduct in Appendix \ref{App_density_to_freq}; Thus, simply dividing equation \eqref{nD_Ito_SDE} by the (now constant) total population size provides the complete dynamics of the system in frequency space: Note that the directional terms in equation \eqref{nD_Ito_SDE} depend only on $\mathbf{A}^-$, which in turn depends only on the fitness $w_i$ and the mutation terms, and this system thus has no noise-induced selection.} such as the neutral Wright-Fisher or Moran models, suggesting that working with such constant population frameworks is not sufficient to accurately capture the dynamics of real populations. Some previous theoretical studies have pointed out that approximating the population size of fluctuating populations through a constant `effective population size' obtained as the harmonic mean of population size over time is not always valid~\citep{sjodin_meaning_2005,parsons_consequences_2010,iizuka_effective_2010,abu_awad_effects_2018,kuosmanen_turnover_2022}. Recent experimental evolution studies have also directly shown that the harmonic mean of population size need not be a good proxy to capture (even approximate) evolutionary dynamics in fluctuating populations~\citep{chavhan_larger_2019}. In our case, noise-induced selection cannot be captured by a constant effective population size, since noise-induced selection involves the turnover $\tau_i$ and is thus a property of the \emph{type} and not a property of the population (and thus each type demands its own effective population size). My results thus lend further support to one of the general messages of these previous studies - Approximating fluctuating populations via a simpler constant (effective) population size may inadvertently remove important evolutionary properties of the systems under study and is thus not always a good idea.

>>>>

The equations of section \ref{sec_fun_theorems} generically predict a directional evolutionary force acting on variation in per-capita turnover rate $\tau$ that I call `noise-induced selection' (following previous studies such as~\cite{constable_demographic_2016} and~\cite{week_white_2021}). Noise-induced selection is only seen in finite populations, is seen whenever there is differential turnover rate $\tau$ in the system, and arises due to different types of individuals in the population experiencing a different number of stochastic events (birth and death) in a given time interval. Several specific finite population models, especially in social evolution, illustrate that evolution in finite population can proceed in a direction different from that predicted by the infinite population limit~\citep{parsons_consequences_2010,melbinger_evolutionary_2010, houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,debarre_evolutionary_2016, behar_fluctuations-induced_2016, constable_demographic_2016,abu_awad_effects_2018,parsons_pathogen_2018,mcavoy_public_2018,mcleod_social_2019}. Indeed, in some stochastic models, the outcome of finite population models can be exactly opposite to that of the infinite population limit, a phenomenon sometimes referred to as `reversing the direction of deterministic selection'~\citep{constable_demographic_2016,mcleod_social_2019}. Chapter \ref{chap_unification} explains how the  results of all these studies can be explained using noise-induced selection. Though similar effects have been recognized in several specific models of life-history evolution~\citep{gillespie_natural_1974,kuosmanen_turnover_2022}, social evolution~\citep{houchmandzadeh_selection_2012,houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,constable_demographic_2016,mcleod_social_2019}, and epidemiology~\citep{parsons_pathogen_2018,day_price_2020}, this thesis is, to the best of my knowledge, the first time it has been explicitly systematically derived and formulated in a general, unifying, `model-independent' evolutionary framework.

Most recently, a study showed (among other things) that in a very broad class of competition models of population ecology, finite populations can directionally evolve in a manner opposite to the direction of evolution predicted in the infinite population limit, essentially due to the effects of noise-induced selection~\citep{mazzolini_universality_2022}. My thesis generalizes this particular result of~\cite{mazzolini_universality_2022} to models with arbitrary interaction types and presents the relevant equations in a form that highlights connections with standard results in evolutionary theory such as the Price equation and Fisher's fundamental theorem (see section \ref{sec_fun_theorems}). It is worth noting that the existence of noise-induced selection directly implies (from equation \eqref{nD_stochastic_RM} or \eqref{nD_stochastic_Price}) that evolution is not expected to maximize fitness in finite populations even if fitness is entirely frequency independent as long as there is some (heritable) variation in the turnover rates $\tau_i$, further underscoring the now well-appreciated point that the view of evolution as `climbing a hill' on a fitness landscape and thereby maximizing fitness is rather limited~\citep{grodwohl_theory_2017}.

The equations of section \ref{sec_fun_theorems} also imply that for the evolution of a trait to be \emph{truly} neutral in finite populations (in the sense of all $m$ types in a system having equal fixation/extinction probability if we start with an initial state in which every type has frequency $1/m$), it is not sufficient for the trait in question to be neutral with respect to fitness $w$. Instead, we also require the trait to be neutral with respect to turnover rate $\tau$. Indeed, in ecological models, previous work in the `quasi-neutral' regime has shown that in finite, fluctuating populations, the equal growth rate of types is not sufficient to ensure equal fixation probabilities and that there is a slight biasing for types with lower turnover rates, sometimes interpreted as a selection `for longevity'~\citep{lin_features_2012, oliveira_advantage_2017,balasekaran_quasi-neutral_2022}. In models of evolutionary game theory in fluctuating finite populations, individuals with lower death rates have higher fixation probability even when growth rates are equalized~\citep{huang_stochastic_2015, czuppon_fixation_2018}. Similarly, models of cell cycle dynamics find that selection favors cell types that periodically arrest their cell cycle relative to non-arresting cells even when their growth rates are equal~\citep{wodarz_effect_2017}. In the language of the birth-death formalism I develop in this thesis, all of these studies equalize the growth rates $w$ of competing types but allow the turnover to vary (by reducing the birth rate through arresting the cell cycle, for example), thus allowing noise-induced selection for reduced turnover to operate in the system. My derivations show analytically that such deviations from neutrality in finite populations are a generic phenomenon explained by noise-induced selection and should be expected whenever there is variation in turnover rates. Selection on turnover rates also leads to insights on life-history evolution, and these insights have been extensively reported in a recent pre-print that independently arrives at my equations for type frequencies (equation \eqref{nD_stochastic_RM}) and the change of mean fitness and turnover in the population (equations \eqref{nD_stochastic_Fisher} and \eqref{nD_stochastic_Fisher_turnover} respectively) using certain discrete time stochastic processes and their approximation via techniques reminiscent of numerical stochastic integration~\citep{kuosmanen_turnover_2022}.

On the practical side, the existence of noise-induced selection implies that simulation studies working with evolutionary individual-based or agent-based models should be careful about whether interaction effects are incorporated into birth rates or death rates since this seemingly arbitrary choice can have unintended consequences due to noise-induced selection, thus potentially biasing results~\citep{mcleod_social_2019,kuosmanen_turnover_2022}. My results also indicate that measuring the growth rate of populations is not, in general, sufficient for accurate prediction/inference of future trajectories of the relative abundance of a species (or phenotype, allele, etc.) from empirical data even in completely controlled environments. The growth rate $w_i = b^{\textrm{(ind)}}_{i} - d^{\textrm{(ind)}}_{i}$ of a species $i$ only specifies the difference between its per-capita birth and death rates. In contrast, the complete stochastic dynamics also depend on the total turnover $\tau_i = b^{\textrm{(ind)}}_{i} + d^{\textrm{(ind)}}_{i}$ (\emph{i.e.} the sum of the per-capita birth and death rates). This discrepancy can lead to systematic deviations from the trajectories predicted using the growth rate alone, especially in smaller populations. However, since the strength of noise-induced selection scales inversely with the total population size, we can expect such deviations to quickly become too small to detect in medium to large sized populations, though this is purely due to limits on the precision of experimental measurements rather than due to any theoretical limitations.

@@

Lastly, noise-induced selection is particular to fluctuating populations and does not occur in models with fixed population sizes\footnote{If $\sum_j x_j$ is a constant, the map $x_i \to x_i/\sum_j x_j$ becomes a linear map and we no longer need It\^o's formula to move from densities to frequencies in the derivation I conduct in Appendix \ref{App_density_to_freq}; Thus, simply dividing equation \eqref{nD_Ito_SDE} by the (now constant) total population size provides the complete dynamics of the system in frequency space: Note that the directional terms in equation \eqref{nD_Ito_SDE} depend only on $\mathbf{A}^-$, which in turn depends only on the fitness $w_i$ and the mutation terms, and this system thus has no noise-induced selection.} such as the neutral Wright-Fisher or Moran models, suggesting that working with such constant population frameworks is not sufficient to accurately capture the dynamics of real populations. Several theoretical studies have pointed out that the empirically common practice of approximating the size of fluctuating populations through a constant `effective population size' (often obtained as the harmonic mean of population size over time) is of limited applicability~\citep{gillespie_natural_1974,sjodin_meaning_2005,parsons_consequences_2010,iizuka_effective_2010,abu_awad_effects_2018,kuosmanen_turnover_2022}. Recent experimental evolution studies have also directly shown that the harmonic mean of population size need not be a good proxy for capturing and/or predicting evolutionary dynamics in fluctuating populations~\citep{chavhan_larger_2019}. In our case, noise-induced selection cannot be captured by any constant effective population size, harmonic mean or otherwise, since noise-induced selection effects also depend on the turnover $\tau_i$ of each type and are thus a property of each \emph{type} in a fluctuating population (and not merely a function of the population size as a whole), meaning that type would demand its own `effective' population size for the entire population. My results thus lend further support to one of the general messages of these previous studies - Approximating fluctuating populations via a simpler constant (effective) population size may inadvertently remove important evolutionary properties of the systems under study and is thus not always a good idea.
